\hrule\hrule\bigskip\bigskip A simple and efficient algorithm for computing approximate Nash equilibria in two-person zero-sum games with imcomplete information

### 2015-06-07 nips2015.tex
more

\section{Introduction}
\label{sec:intro}
\paragraph{\textbf{Context.}}
A game-theoretic approach to playing games strategically optimally
consists in computing Nash-equilibria (infact, approximations thereof)
offline, and playing one's part of the equilibrium online. This
technique is the driving-force behind solution concepts like CFR
(\textbf{TODO: add Z.'s ref first!!!}) \cite{lanctot2009monte},
$\text{CFR}^{+}$ \textbf{TODO: add ref!!!}, and other variants, which
have recently had profound success in Poker. However, solving games
for equilibria remains a mathematical and computational challenge,
especially in sequential games with imperfect information. This paper
proposes a simple and fast algorithm for solving for such equilibria
approximately, in a sense which will be made clear shortly.

\paragraph{\textbf{Notation and terminology.}}
Let $m$ and $n$ be positive integers.
$\mathbb{R}^{n}_+ := \{z \in \mathbb{R}^{n}\text{ }|\text{ } z \geq
0\}$ is the nonnegative $n$-dimensional \textit{orthant}.  The
components of a vector $z \in \mathbb{R}^n$ will be denoted $z_0$,
$z_1$, ..., $z_{n-1}$ (indexing begins from $0$, and not $1$). The
notation ``$z \ge 0$'' means that all the components of $z$ are
nonnegative. For a vector $z = (z_0, z_1, ..., z_{n-1}) \in
\mathbb{R}^n$, %% $\|z\|$ denotes the $2$-\textit{norm} of $z$ defined
%% by $\|z\| := \sqrt{z^Tz} = \sqrt{\sum_{j=0}^{n - 1}{z_j^2}}$.
$(z)_+:=\text{max}(0, z) \in \mathbb{R}^{n}_+$ is the point-wise
maximum of $z$ with $0$.
%% For example, $((-2, \pi))_+ = (max(-2, 0),
%% max(\pi, 0)) = (0, \pi)$.
The operator $(.)_+$ is the well-known
(multi-dimensional) \textit{ramp} function.
$\mathbb{R}^{m \times n}$ denotes the space of all $m$-by-$n$ real
matrices, $I_m$ denotes the $m$-by-$m$ identity matrix, and $0_{m,n}$
denotes the $m$-by-$n$ matrix of zeros.
Given a matrix $A \in \mathbb{R}^{m \times n}$, its \textit{spectral
  norm}, denoted $\|A\|$, is
 defined to be the largest \textit{singular value} of $A$, i.e the
 largest \textit{eigenvalue} of $A^TA$ (or equivalently, of $AA^T$).

Given a subset $C$ of $\mathbb{R}^n$,
$i_C$ denotes its \textit{indicator function} defined by
\begin{eqnarray}
  i_C(x) = 0, \text{ if } x \in C;\hspace{.3em} i_c(x) = \infty\text{
    otherwise.}
\end{eqnarray}

Let $f : \mathbb{R}^n \rightarrow (-\infty, +\infty]$ be a convex
  function.
The \textit{effective domain} of $f$, denoted $\dom(f)$ is defined as
\begin{eqnarray}
  \dom(f) := \{x \in \mathbb{R}^n | f(x) < +\infty\}.
\end{eqnarray}
 If $\dom(f) \ne \emptyset$ then we say $f$ is \textit{proper}.
%% \textit{proper convex lower semi-continous function}
%% (\textit{p.c.l.s.c} for short).
%% The \textit{Fenchel-Legendre transform} of $f$ is the function $f^*:
%% \mathbb{R}^n \rightarrow (-\infty, +\infty]$ defined by $f^*(x) \equiv
%%   \underset{z \in \mathbb{R}^n}{\text{max}}\text{}z^Tx - f(z)$.
Given $\tau > 0$, the function
\begin{eqnarray}
  \prox_{\tau f}: \mathbb{R}^n \rightarrow \mathbb{R}^n,\hspace{.5em}x
  \mapsto \underset{z \in \mathbb{R}^n}{\text{min }}\frac{1}{2}\|z
  - x\|^2 + \tau f(z),
\end{eqnarray}

@@

\paragraph{\textbf{Statement of the problem.}}
The sequence-form for two-person zero-sum games was introduced in
\cite{koller1992complexity}, and the theory was further developed in
\cite{koller1994fast,von1996efficient,vonequilibrium}, where it was
established that for sequential two-person zero-sum games with
imcomplete information and perfect recall, there exist sparse matrices
$A \in \mathbb{R}^{n_1
  \times n_2}$, $E_1 \in \mathbb{R}^{l_1 \times n_1}$, $E_2 \in
\mathbb{R}^{l_2 \times n_2}$, and vectors $e_1 \in \mathbb{R}^{l_1},
e_2 \in \mathbb{R}^{l_2}$ such that $n_1$, $n_2$, $l_1$, and $l_2$ are
all linear in the size of the game tree (number of states in the game)
and such that Nash equilibria correspond to pairs $(x, y)$ of
\textit{realization plans} which solve the primal LCP
\begin{equation}
  \begin{aligned}
     \text{minimize }e_1^Tp, \hspace{.5em}\text{ subject to }(y,p) \in
     \mathbb{R}^{n_2} \times \mathbb{R}^{l_1}, -Ay +
    E_1^Tp \geq 0, E_2y = e_2, y \ge 0.
  \end{aligned}
  \label{eq:primal_pb}
\end{equation}

@@

\paragraph{\textbf{Our contribution.}}
We develop a $\mathcal{O}(1/\text{min}(\rho,\epsilon))$ primal-dual
algorithm for computing Nash $(\rho, \epsilon$)-equilibria (in a sense
to be made clear later) for
sequential two-person zero-sum games with imcomplete information and
perfect recall.

@@

\begin{definition}[$\epsilon$-subgradient]
Given a scalar $\epsilon > 0$ and a function $f:X \subseteq \mathbb{R}^n
\rightarrow [-\infty,+\infty]$, the $\epsilon$-subgradient of $f$ the
set-valued function
\begin{eqnarray}
\partial_\epsilon f: X \mapsto 2^X, \hspace{.5em} x \mapsto \{h\in X |
f(z) + \epsilon \ge f(x) + \langle h, z - x\rangle \forall z \in X\}.
\end{eqnarray}
\end{definition}

@@

Algorithm \ref{Tab:algo} is then obtained from \cite[Algorithm
  T-BD]{he2013accelerating}, applied on the GSP
\eqref{eq:unconstrained_pb} with the choice of parameters: $\sigma = 1
\in (0, 1]$, $\sigma_x = \sigma_y = 0 \in [0, \sigma)$,
    $\lambda_{xy} := \frac{\sqrt{(\sigma^2 - \sigma_x^2)(\sigma^2 -
        \sigma_y^2)}}{\sigma L_{xy}} = \sigma / \|K\| =
    1/\|K\|$, and $\lambda = \lambda_{xy} \in (0,
    \lambda_{xy}]$. The convergence result follows immediately from
  \cite[Theorem 4.2]{he2013accelerating}.
  \qed
\end{proof}

>>>>

\section{Introduction}
\label{sec:intro}
A game-theoretic approach to playing games strategically optimally
consists in computing Nash-equilibria (infact, approximations thereof)
offline, and playing one's part of the equilibrium online. This
technique is the driving-force behind solution concepts like CFR
(\textbf{TODO: add Z.'s ref first!!!}) \cite{lanctot2009monte},
$\text{CFR}^{+}$ \textbf{TODO: add ref!!!}, and other variants, which
have recently had profound success in Poker. However, solving games
for equilibria remains a mathematical and computational challenge,
especially in sequential games with imperfect information. This paper
proposes a simple and fast algorithm for solving for such equilibria
approximately, in a sense which will be made clear shortly.

\subsection{Notation and terminology}
\paragraph{\textbf{General notions.}}
Let $m$ and $n$ be positive integers.
$\mathbb{R}^{n}_+ := \{z \in \mathbb{R}^{n}\text{ }|\text{ } z \geq
0\}$ is the nonnegative $n$-dimensional \textit{orthant}.  The
components of a vector $z \in \mathbb{R}^n$ will be denoted $z_0$,
$z_1$, ..., $z_{n-1}$ (indexing begins from $0$, and not $1$). The
notation ``$z \ge 0$'' means that all the components of $z$ are
nonnegative. For a vector $z = (z_0, z_1, ..., z_{n-1}) \in
\mathbb{R}^n$, %% $\|z\|$ denotes the $2$-\textit{norm} of $z$ defined
%% by $\|z\| := \sqrt{z^Tz} = \sqrt{\sum_{j=0}^{n - 1}{z_j^2}}$.
$(z)_+:=\text{max}(0, z) \in \mathbb{R}^{n}_+$ is the point-wise
maximum of $z$ with $0$.
%% For example, $((-2, \pi))_+ = (max(-2, 0),
%% max(\pi, 0)) = (0, \pi)$.
The operator $(.)_+$ is the well-known
(multi-dimensional) \textit{ramp} function.
$\mathbb{R}^{m \times n}$ denotes the space of all $m$-by-$n$ real
matrices, $I_m$ denotes the $m$-by-$m$ identity matrix, and $0_{m,n}$
denotes the $m$-by-$n$ matrix of zeros.
Given a matrix $A \in \mathbb{R}^{m \times n}$, its \textit{spectral
  norm}, denoted $\|A\|$, is
 defined to be the largest \textit{singular value} of $A$, i.e the
 largest \textit{eigenvalue} of $A^TA$ (or equivalently, of $AA^T$).

\paragraph{\textbf{Convex analysis.}}
Given a subset $C$ of $\mathbb{R}^n$,
$i_C$ denotes its \textit{indicator function} defined by
\begin{eqnarray}
  i_C(x) = 0, \text{ if } x \in C;\hspace{.3em} i_c(x) = \infty\text{
    otherwise.}
\end{eqnarray}

Let $f : \mathbb{R}^n \rightarrow (-\infty, +\infty]$ be a convex
  function.
The \textit{effective domain} of $f$, denoted $\dom(f)$ is defined as
\begin{eqnarray}
  \dom(f) := \{x \in \mathbb{R}^n | f(x) < +\infty\}.
\end{eqnarray}
 If $\dom(f) \ne \emptyset$ then we say $f$ is \textit{proper}.
%% \textit{proper convex lower semi-continous function}
%% (\textit{p.c.l.s.c} for short).
%% The \textit{Fenchel-Legendre transform} of $f$ is the function $f^*:
%% \mathbb{R}^n \rightarrow (-\infty, +\infty]$ defined by $f^*(x) \equiv
%%   \underset{z \in \mathbb{R}^n}{\text{max}}\text{}z^Tx - f(z)$.
The \textit{subgradient} of $f$ is the set-valued function
\begin{eqnarray}
\partial f: X \mapsto 2^X, \hspace{.5em} x \mapsto \{h\in X |
f(z)  \ge f(x) + \langle h, z - x\rangle \forall z \in X\}.
\end{eqnarray}
Of course $\partial f(x)$ reduces to the singleton $\{\nabla f(x)\}$
in case $f$ is differentiable at $x$. Given $\tau > 0$, the function
\begin{eqnarray}
  \prox_{\tau f}: \mathbb{R}^n \rightarrow \mathbb{R}^n,\hspace{.5em}x
  \mapsto \underset{z \in \mathbb{R}^n}{\text{min }}\frac{1}{2}\|z
  - x\|^2 + \tau f(z),
\end{eqnarray}

@@

\subsection{Statement of the problem}
The sequence-form for two-person zero-sum games was introduced in
\cite{koller1992complexity}, and the theory was further developed in
\cite{koller1994fast,von1996efficient,vonequilibrium}, where it was
established that for sequential two-person zero-sum games with
imcomplete information and perfect recall, there exist sparse matrices
$A \in \mathbb{R}^{n_1
  \times n_2}$, $E_1 \in \mathbb{R}^{l_1 \times n_1}$, $E_2 \in
\mathbb{R}^{l_2 \times n_2}$, and vectors $e_1 \in \mathbb{R}^{l_1},
e_2 \in \mathbb{R}^{l_2}$ such that $n_1$, $n_2$, $l_1$, and $l_2$ are
all linear in the size of the game tree (number of states in the game)
and such that Nash equilibria correspond to pairs $(x, y)$ of
\textit{realization plans} which solve the primal LCP
\begin{equation}
  \begin{aligned}
     \text{minimize }e_1^Tp, \hspace{.5em}\text{ subject to }(y,p) \in
     \mathbb{R}^{n_2} \times \mathbb{R}^{l_1}, -Ay +
    E_1^Tp \geq 0, E_2y = e_2, y \ge 0.
  \end{aligned}
  \label{eq:primal_pb}
\end{equation}

@@

\subsection{Our contribution}
We develop a $\mathcal{O}(1/\text{min}(\rho,\epsilon))$ primal-dual
algorithm (Algorithm \ref{Tab:algo}) for computing Nash $(\rho,
  \epsilon$)-equilibria (in a sense to be made clear later) for
  sequential two-person zero-sum games with imcomplete information and
  perfect recall.

@@

\begin{definition}[$\epsilon$-subgradient]
Given a scalar $\epsilon > 0$ and a function $f:X \subseteq \mathbb{R}^n
\rightarrow [-\infty,+\infty]$, the $\epsilon$-enlarged subgradient (or
$\epsilon$-subgradient, for short) of $f$ the set-valued function
\begin{eqnarray}
\partial_\epsilon f: X \mapsto 2^X, \hspace{.5em} x \mapsto \{h\in X |
f(z) + \epsilon \ge f(x) + \langle h, z - x\rangle \forall z \in X\}.
\end{eqnarray}
\end{definition}

@@

Algorithm \ref{Tab:algo} is then obtained from \cite[Algorithm
  T-BD]{he2013accelerating}, applied on the GSP
\eqref{eq:unconstrained_pb} with the choice of parameters: $\sigma = 1
\in (0, 1]$, $\sigma_x = \sigma_y = 0 \in [0, \sigma)$,
    $\lambda_{xy} := \frac{1}{\sigma L_{xy}}\sqrt{(\sigma^2 -
        \sigma_x^2)(\sigma^2 - \sigma_y^2)} = \sigma / \|K\| =
      1/\|K\|$, and $\lambda = \lambda_{xy} \in (0,
      \lambda_{xy}]$. The convergence result follows immediately from
  \cite[Theorem 4.2]{he2013accelerating}.
  \qed
\end{proof}
